Fractional Decomposition Tree Algorithm: A tool for studying the integrality gap of Integer Programs

### 2020-08-07 FDT2020/Introduction.tex
Just responding to the latest comments

%\arash{I added this figure \ref{fig:CVpoint} but maybe we can put it later in the experiments section to make the introduction shorter?} \cindy{Yes that would be fine, if the reader could also use that figure then.  They could page forward to see the figure when reading the introduction.}
\arash{I changed the paragraph below a little bit, can you check?}
Carr and Vempala \cite{Carr2004} showed that $g(\2ec)$ is achieved for instances where the optimal solution to $\min_{x\in \subtour(G)}cx$ is a Carr-Vempala point. It has be conjectured that $g(\2ec)\leq \frac{6}{5}$ (see \cite{alexander2006integrality,boydlegault,hn19}, but the only known upper bound on $g(\2ec)$ stands at $\frac{3}{2}$ \cite{wolsey}. We show that the integrality gap is at most $\frac{6}{5}$ for Carr-Vempala points with at most 12 vertices on the cycle formed by the fractional edges, thereby confirming the conjecture for these instances.
\cindy{This sounds like a contribution of the paper.} \arash{Yes, particlarly since a similar research question have been studied by \cite{alexander2006integrality} and \cite{TSPcompute}}
Note that the number of vertices in these instances can be arbitrarily high since the paths of edges with $x$-value 1 can be arbitrarily long.

>>>>

%\arash{I added this figure \ref{fig:CVpoint} but maybe we can put it later in the experiments section to make the introduction shorter?} \cindy{Yes that would be fine, if the reader could also use that figure then.  They could page forward to see the figure when reading the introduction.}
\arash{I changed the paragraph below a little bit, can you check?} \cindy{Looks good. I just changed one sentence to the active voice.  If you're OK with that, you can delete these 2 comments. Thanks!}
Carr and Vempala \cite{Carr2004} showed that $g(\2ec)$ is achieved for instances where the optimal solution to $\min_{x\in \subtour(G)}cx$ is a Carr-Vempala point. Multiple groups of researchers have conjectured that $g(\2ec)\leq \frac{6}{5}$ (see \cite{alexander2006integrality,boydlegault,hn19}, but the only known upper bound on $g(\2ec)$ stands at $\frac{3}{2}$ \cite{wolsey}. We show that the integrality gap is at most $\frac{6}{5}$ for Carr-Vempala points with at most 12 vertices on the cycle formed by the fractional edges, thereby confirming the conjecture for these instances.
\cindy{This sounds like a contribution of the paper.} \arash{Yes, particlarly since a similar research question have been studied by \cite{alexander2006integrality} and \cite{TSPcompute}}
Note that the number of vertices in these instances can be arbitrarily high since the paths of edges with $x$-value 1 can be arbitrarily long.

### 2020-08-07 FDT2020/main-section.tex
Just responding to the latest comments

Consider an instance $I=(A,b)$ of the IP formulation. Define sets $S(I)$ and $P(I)$ as in (\ref{S}) and (\ref{P}), respectively. Assume $S(I)\subseteq \{0,1\}^n$ and $P(I)\subseteq [0,1]^n$. For simplicity in the notation we assume an instance $I$ and denote $P(I),S(I),$ and $g(I)$ by $P$, $S$, and $g$ for this section and the next section. Also, for both sections let $x^*$ be the optimal solution to the LP formulation and assume $t=|\spp(x^*)|$. Without loss of generality we can assume $x^*_i = 0$ for $i=t+1,\ldots,n$.
\cindy{TODO: Is this something that should be clear to the reader with what they know at this point? Does it need an explanation? Does it need an explanation at the end of this section or the next?}\arash{this is just a relabeling of the indices.} \cindy{Sorry.  I didn't understand that you meant the solution has that property. Right? The vector $x$ isn't defined in this paragraph and it's not in the next lemma either. If $x$ is the output of DomToIP, then we should say that.  ``Without loss of generality the solution $x$ returned by DomToIP has ...''}\arash{I see your point, please see if the change made it better}

In this section we prove Theorem \ref{domtoIP}. In fact, we prove a stronger result. 
\begin{lemma}\label{domlemma}
	Given $\tilde{x}\in \dom(P)$ and $\tilde{x}\in \{0,1\}^n$, there is an algorithm (the DomToIP algorithm) that finds $\bar{x}\in S$ in polynomial time, such that $\bar{x}\leq \tilde{x}$.\end{lemma}
Notice that Lemma \ref{domlemma} implies Theorem \ref{domtoIP}, since it is easy to obtain an integer point in $\dom(P)$: numerically rounding up $x^*$ (or any fractional point in $P$) gives us a point in $\dom(P)$. Hence, we can assume in the proof below that $\tilde{x}_i= 0$ for $i=t+1,\ldots,n$. 
\arash{I also made a change based on the change above}
\subsection{Proof of Lemma \ref{domlemma}: The DomToIP Algorithm}

@@

\section{FDT on Binary IPs}
\label{sec:binaryfdt}
\arash{I made a small change here based on our discussion in the bottom of page 8}
Recall that $x^*$ was the optimal solution of minimizing a cost function $cx$ over set $P$, which provides a lower bound on $\min_{(x,y)\in S(I)} cx$.  In this section, we prove Theorem \ref{binaryFDT} by describing the Fractional Decomposition Tree (FDT) algorithm. We also remark that if $g(I)=1$, then the algorithm will give an exact decomposition of any feasible solution.

@@

\begin{figure}[h!]
\begin{subfigure}{.5\textwidth}
\centering
	\includegraphics[width=8cm,scale=1]{"fpvsfdt".png}
	\caption{Dive FDT vs feasbility pump on the instances of PACE 2019 \cite{PACE} with 200 vertices.}
	\label{fpvsfdt}
	\end{subfigure}
	$\quad\;$
	\begin{subfigure}{.5\textwidth}
	\centering
	\includegraphics[width=8cm,scale=1]{"fjvsfdt".png}
%	\caption{FDT vs the 2-approximation for TAP\cite{FJ81} on randomly generated extreme points of the cut-LP.}
	\caption{TAP on our random instances: FDT run on LP optimal vs the 2-approximation for TAP\cite{FJ81}.}
	\label{fjvsfdt}
	\end{subfigure}
	\caption{Computational experiments with the FDT algorithm}
	\label{fdtcomp}
\end{figure}
\paragraph{FDT on randomly generated instances of TAP.}
Recall that in the tree-augmentation problem (TAP) we are given a tree $T=(V,E)$, a set of non-tree links $L$ between vertices in $V$ and costs $c\in \mathbb{R}^{L}_{\geq 0}$. A feasible augmentation is $L'\subseteq L$ such that $T+L'$ is 2-edge-connected. In TAP we wish to find the minimum-cost feasible augmentation. The integrality gap of the cut LP for TAP (given in (\ref{eq:cutLP})) is
\begin{equation*}
g(\tap) = \max_{c\in \mathbb{R}^L_{\geq 0}} \frac{\min_{x\in\tap(T,L)} cx}{\min_{x\in\cut(T,L)} cx},
\end{equation*}  
where $\tap(T,L)$ is the feasible set for the IP and $\cut(T,L)$ is the feasible set for the cut LP (relaxation).
\arash{Did you intentionally change the TAP and CUT module? They print as italic now as opposed to how they look in $\tap(T,L)$ and $\cut(T,L)$}
\cindy{Did you mean macro?  No, I didn't intentionally change it.  I just hadn't noticed you had a macro for those two and was in a rush.  At least for the previous section, I changed to the macros. I don't immediately see other places where it's used in math mode, and therefore needs the macro.}
We know $\frac{3}{2}\leq g(\tap)\leq 2$~\cite{FJ81,32gaptap}. Notice that $\min_{x\in \tap(T,L)}cx$ is a binary IP.

>>>>

Consider an instance $I=(A,b)$ of the IP formulation. Define sets $S(I)$ and $P(I)$ as in (\ref{S}) and (\ref{P}), respectively. Assume $S(I)\subseteq \{0,1\}^n$ and $P(I)\subseteq [0,1]^n$. For simplicity in the notation we assume an instance $I$ and denote $P(I),S(I),$ and $g(I)$ by $P$, $S$, and $g$ for this section and the next section. Also, for both sections let $x^*$ be the optimal solution to the LP formulation and assume $t=|\spp(x^*)|$. Without loss of generality we can assume $x^*_i = 0$ for $i=t+1,\ldots,n$.
\cindy{TODO: Is this something that should be clear to the reader with what they know at this point? Does it need an explanation? Does it need an explanation at the end of this section or the next?}\arash{this is just a relabeling of the indices.} \cindy{Sorry.  I didn't understand that you meant the solution has that property. Right? The vector $x$ isn't defined in this paragraph and it's not in the next lemma either. If $x$ is the output of DomToIP, then we should say that.  ``Without loss of generality the solution $x$ returned by DomToIP has ...''}\arash{I see your point, please see if the change made it better} \cindy{Yes, thanks.  We can delete this conversation after you see this response.}

In this section we prove Theorem \ref{domtoIP}. In fact, we prove a stronger result. 
\begin{lemma}\label{domlemma}
	Given $\tilde{x}\in \dom(P)$ and $\tilde{x}\in \{0,1\}^n$, there is an algorithm (the DomToIP algorithm) that finds $\bar{x}\in S$ in polynomial time, such that $\bar{x}\leq \tilde{x}$.\end{lemma}
Notice that Lemma \ref{domlemma} implies Theorem \ref{domtoIP}, since it is easy to obtain an integer point in $\dom(P)$: numerically rounding up $x^*$ (or any fractional point in $P$) gives us a point in $\dom(P)$. Hence, we can assume in the proof below that $\tilde{x}_i= 0$ for $i=t+1,\ldots,n$. 
\arash{I also made a change based on the change above}\cindy{Thanks!}
\subsection{Proof of Lemma \ref{domlemma}: The DomToIP Algorithm}

@@

\section{FDT on Binary IPs}
\label{sec:binaryfdt}
\arash{I made a small change here based on our discussion in the bottom of page 8}\cindy{Thanks!}
Recall that $x^*$ was the optimal solution of minimizing a cost function $cx$ over set $P$, which provides a lower bound on $\min_{(x,y)\in S(I)} cx$.  In this section, we prove Theorem \ref{binaryFDT} by describing the Fractional Decomposition Tree (FDT) algorithm. We also remark that if $g(I)=1$, then the algorithm will give an exact decomposition of any feasible solution.

@@

\begin{figure}[h!]
\begin{subfigure}{.5\textwidth}
\centering
	\includegraphics[width=8cm,scale=1]{"fpvsfdt".png}
	\caption{Dive FDT vs feasbility pump on the instances of PACE 2019 \cite{PACE} with 200 vertices.}
	\label{fpvsfdt}
	\end{subfigure}
	$\quad\;$
	\begin{subfigure}{.5\textwidth}
	\centering
	\includegraphics[width=8cm,scale=1]{"fjvsfdt".png}
%	\caption{FDT vs the 2-approximation for TAP\cite{FJ81} on randomly generated extreme points of the cut-LP.}
	\caption{TAP on our random instances: FDT run on LP optimal vs the 2-approximation for TAP\cite{FJ81}.}
	\label{fjvsfdt}
	\end{subfigure}
	\caption{Computational experiments with the FDT algorithm}
	\label{fdtcomp}
\end{figure}
\paragraph{FDT on randomly generated instances of TAP.}
Recall that in the tree-augmentation problem (TAP) we are given a tree $T=(V,E)$, a set of non-tree links $L$ between vertices in $V$ and costs $c\in \mathbb{R}^{L}_{\geq 0}$. A feasible augmentation is $L'\subseteq L$ such that $T+L'$ is 2-edge-connected. In TAP we wish to find the minimum-cost feasible augmentation. The integrality gap of the cut LP for TAP (given in (\ref{eq:cutLP})) is
\begin{equation*}
g(\tap) = \max_{c\in \mathbb{R}^L_{\geq 0}} \frac{\min_{x\in\tap(T,L)} cx}{\min_{x\in\cut(T,L)} cx},
\end{equation*}  
where $\tap(T,L)$ is the feasible set for the IP and $\cut(T,L)$ is the feasible set for the cut LP (relaxation).
%\arash{Did you intentionally change the TAP and CUT module? They print as italic now as opposed to how they look in $\tap(T,L)$ and $\cut(T,L)$}
%\cindy{Did you mean macro?  No, I didn't intentionally change it.  I just hadn't noticed you had a macro for those two and was in a rush.  At least for the previous section, I changed to the macros. I don't immediately see other places where it's used in math mode, and therefore needs the macro.}
We know $\frac{3}{2}\leq g(\tap)\leq 2$~\cite{FJ81,32gaptap}. Notice that $\min_{x\in \tap(T,L)}cx$ is a binary IP.
